\section{Introduction}
Eye-tracking is invaluable at explaining how people perceive, solve visual tasks, and use interfaces ~\cite{duchowski2002breadth}. However, much of the technology's full potential is untapped, as most eye-tracking studies explore key-hole scenarios involving static images, simple tasks, and no possibility for subjects to interact with visual content. This is at least in part because gazes are traditionally analyzed as pixel coordinates of evaluated visual frames. Relating these gaze coordinates back to the semantic content of each visual frame viewed in an experiment (i.e., determining whether a gaze coordinate matches a certain object in the image) involves extensive manual input, and is prohibitively time-consuming for complex experiments.
 
Recently, there is increasing interest in methods that look at what users are looking rather than where they are looking~\cite{alam15analyzing, sundstedt2013visual, bernhard2014gaze}. Such methods are applicable to visual content that is computer generated. Because positions and shapes of objects shown on the screen are known during rendering, gaze coordinates can be mapped automatically and in real time to a visualization's semantic content.  

Such methods produce temporal sequences of data-objects that subjects view during an experiment (e.g., nodes in a network, 3D objects in a scene). We will refer to these objects as data-of-interest (DOI), a name used by Alam et al. to underline the analogy with areas of interest (AOI). Since DOIs are a subset of the data underlying an experiment, they are richly described by attributes that are derived from that data (e.g., labels, types of objects, quantitative attributes). Moreover, DOI methods capture subjects' interest in DOIs irrespective of the visual context in which these are observed. For example, in a node-link diagram, a user's interest in a node will be tracked regardless of whether that node moves as a result of interactions, or whether the visualization as a whole is zoomed or panned. Such analyses are thus immune to a subject's interactions with a system, meaning that once a visualization is instrumented, data can be captured without additional effort from interactive systems over extended periods of time~\cite{alam15analyzing}. Moreover, as long as subjects explore the same data-space, their results will be comparable in that space, even if they do not see the same rendered views. 

These properties mean that DOI methods can support research workflows not previously possible~\cite{alam15analyzing, sundstedt2013visual, bernhard2014gaze}. Our paper makes two necessary contributions on our way to applying DOIs methods in practice. 
Our first contribution is to formalize the range of questions that can be asked of DOI data by introducing a novel DOI task taxonomy. Task-taxonomies help drive the development of novel visual encodings and visual analytics designs by formalizing requirements and supporting evaluation ~\cite{amar2005low,okoe2015graphunit}. However, no task taxonomies exist for DOI data, or even the more established AOI analyses. As we will show, this contribution is significant as DOI data can support tasks and workflows significantly different from those that AOIs could be applied for.

Our second contribution is to define and discuss the design space of visual encodings and visual analytics principles that can support DOI workflows in general, and the DOI task taxonomy in particular. New visual designs are necessary because existing methods for AOI analysis do not scale to DOI data, and do not support the types of questions that DOI data can answer. To define a new design space, we start from existing visual methods for AOI analysis, and explore the ways in which they can be adapted to support DOI analyses. We structure our exploration using Yi et al.'s~\cite{yi2007toward} taxonomy of visualization interactions, consider solutions that can support most of the tasks described in our DOI task taxonomy, and discuss the strengths and weaknesses of alternatives.  This analysis represents a baseline for future design studies and systems exploring the use of DOI analyses in practice.


