\begin{acknowledgments}
I express my heartiest gratitude to Dr. Radu Jianu who was my supervisor for the first four years. He kept in touch with me even after leaving the university and invariably helped me in completing the ongoing research. Throughout the years, his enthusiasm, assistance, and foresight gave me the necessary strength to proceed through the doctoral program and finish this dissertation. 

I would like to express my sincere appreciation to my supervisor Dr.  S. S. Iyengar for his help and support during the last years of my doctoral program. I am thankful to my committee members Dr. Finlayson, Dr. Wei Zeng, Dr. Bobadilla, and Dr. Kibria for their valuable recommendations to improve this dissertation. I also thank Dr. Nipesh Pradhananga from the school of construction and Dr. Shahin Vassigh from the department of architecture at FIU for collaborating on two projects which contributed to this dissertation.

I like to thank my family members who supported me over the last five years. My mother Shahin Sultana struggled her whole life to see me successful. This dissertation would not be even possible without her love, patience, and guidance.  I could not thank enough my wife, Nahid Ferdous. She always motivated me to complete my dissertation. I am also thankful to my sister Fahrina Alam for her unconditional support over the years.  I thank my labmate, Dr. Mershack Okoe. We worked together on many projects, and my discussions with him regarding several research problems greatly benefited me in countless scenarios. 
\end{acknowledgments}